---
title: "Data Science Capstone - Milestone Report"
author: "Gaston Dario Napoli"
date: "May 11, 2017"
output: 
    html_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) Introduction

This document corresponds to the ***Milestone Report***, assignment of week 2, [***Data Science Capstone***](https://www.coursera.org/learn/data-science-project) course from [***Coursera***](https://www.coursera.org/). This course, the 10th out of 10 courses comprising the [***Data Science Specialization***](<https://www.coursera.org/specializations/jhu-data-science>) from the [***John Hopkins University***](https://www.jhu.edu/), allows students to create a usable/public data product that can be used to show your skills to potential employers. Projects will be drawn from real-world problems and will be conducted with industry, government, and academic partners.

The ***Data Science Capstone*** project aims to develop a [***Shiny***](https://shiny.rstudio.com) app that takes as input a phrase (multiple words), one clicks submit, and it will predict the next word. In order to achive that, the [***course dataset***](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip) will be used as a training set, as well as NPL techniques will be applied to analyse and build the corresponding predictive model.

## 2) Dataset

Borrowing from the course [dataset information](https://www.coursera.org/learn/data-science-project/supplement/Iimbd/task-0-understanding-the-problem), the data comes from a corpus called ***HC Corpora*** (the original site is not reachable, but an archive of it can be seen at https://web-beta.archive.org/web/20160930083655/http://www.corpora.heliohost.org/aboutcorpus.html).  ***corpora*** are collected from publicly available sources by a web crawler.

As said before, the dataset can be downloaded from https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip, and consists, once being downloaded and uncompressed, of 4 folders corresponding to 4 different languages (german,english, finish and russian), and each folder containing 3 files from 3 different text sources (blogs, news and Twitter). Next, a summary table of each file can be seen (file stats plus **wc** command data):

```{r echo=FALSE}
library(knitr)
library(stringr)

setwd("D:/OneDrive/Documentos/Coursera/Capstone/Material")

# Get the file list
listOfFiles <- dir(recursive = TRUE)

# Get the file stat list from each file
listOfFileInfos <- file.info(listOfFiles)

# Generate three new columns in otrder to comple with 'wc' command execution data
listOfFileInfos$lineCount <- 0
listOfFileInfos$wordCount <- 0
listOfFileInfos$characterCount <- 0

# Complete de file stats with the 'wc' command data
for (file in listOfFiles) {
    wcCommandResults <- str_split(system(paste0("wc ", file), intern = TRUE),  boundary("word"))[[1]]
	
    listOfFileInfos[file,]$lineCount <- wcCommandResults[1]
    listOfFileInfos[file,]$wordCount <- wcCommandResults[2]
    listOfFileInfos[file,]$characterCount <- wcCommandResults[3]
}

# Show a formatted table
kable(listOfFileInfos[,c("size", "ctime", "lineCount", "wordCount", "characterCount")])

```

In the context of the Capstone project, only the english language files will be taken into account.

## Apendix I - Source codes

This document has been generated using [***R Mardown***](http://rmarkdown.rstudio.com/). Its ***.Rmd*** source code that can be found at: <https://github.com/laplata2003/data-science-capstone-week2-milestone-report>.

