---
title: "Data Science Capstone - Milestone Report"
author: "Gaston Dario Napoli"
date: "May 12, 2017"
output: 
    html_document:
        toc: true
---

```{r setup, include=FALSE}
library(knitr)
library(stringr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
```

## 1) Introduction

This document corresponds to the ***Milestone Report***, assignment of week 2, [***Data Science Capstone***](https://www.coursera.org/learn/data-science-project) course from [***Coursera***](https://www.coursera.org/). This course, the 10th out of 10 courses comprising the [***Data Science Specialization***](<https://www.coursera.org/specializations/jhu-data-science>) from the [***John Hopkins University***](https://www.jhu.edu/), allows students to create a usable/public data product that can be used to show your skills to potential employers. Projects will be drawn from real-world problems and will be conducted with industry, government, and academic partners.

The ***Data Science Capstone*** project aims to develop a [***Shiny***](https://shiny.rstudio.com) app that takes as input a phrase (multiple words), one clicks submit, and it will predict the next word. In order to achive that, the [***course dataset***](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip) will be used as a training set, as well as NPL techniques will be applied to analyse and build the corresponding predictive model.

## 2) Dataset
### 2.1) Description

From the [course dataset information](https://www.coursera.org/learn/data-science-project/supplement/Iimbd/task-0-understanding-the-problem), the data comes from a corpus called ***HC Corpora*** (the original site is not reachable, but an archive of it can be seen at https://web-beta.archive.org/web/20160930083655/http://www.corpora.heliohost.org/aboutcorpus.html).  ***Corpora*** are collected from publicly available sources by a web crawler.

As said before, the dataset can be downloaded from https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip, and consists, once being downloaded and uncompressed, of 4 folders corresponding to 4 different languages (german,english, finish and russian), and each folder containing 3 files from 3 different text sources (blogs, news and Twitter):

```{r results='asis'}

# Auxiliary function. It allows to show a list as bullets (use of results='asis' is mandatory)
showBullets <- function(listOfValues) {
    for (i in listOfValues){
        cat("-", i, "\n")
    }
}

# Get the file list
listOfFiles <- dir("HC_Corpora", recursive = TRUE, full.names = TRUE)

# Show the list as bullets
showBullets(listOfFiles)
```


### 2.2) Analysis
Our analysis starts with a summary table including, for each file in the bundle, file stats (size in bytes) and ***wc*** command data (i.e. lines, words and characters counting):

```{r}
# Get the file stat list from each file
listOfFileInfos <- file.info(listOfFiles)

# Generate three new columns in otrder to comple with 'wc' command execution data
listOfFileInfos$lineCount <- 0
listOfFileInfos$wordCount <- 0
listOfFileInfos$characterCount <- 0

# Complete de file stats with the 'wc' command data
for (file in listOfFiles) {
    wcCommandResults <- str_split(system(paste0("wc ", file), intern = TRUE),  boundary("word"))[[1]]
	
    listOfFileInfos[file,]$lineCount <- as.numeric(wcCommandResults[1])
    listOfFileInfos[file,]$wordCount <- as.numeric(wcCommandResults[2])
    listOfFileInfos[file,]$characterCount <- as.numeric(wcCommandResults[3])
}

# Show a formatted table
kable(listOfFileInfos[,c("size", "ctime", "lineCount", "wordCount", "characterCount")])

```

In the context of the Capstone project, only the english language files will be taken into account, that is to say, the files with the ***en_US*** prefix:

```{r results='asis'}

englishFiles <- listOfFiles[startsWith(listOfFiles, "HC_Corpora/en_US")]
showBullets(englishFiles)

```


## Apendix I - Source codes

This document has been generated using [***R Mardown***](http://rmarkdown.rstudio.com/). Its ***.Rmd*** source code that can be found at: <https://github.com/laplata2003/data-science-capstone-week2-milestone-report>.

